Carbonate chemistry at Ny-Ålesund

### 2023-01-10 awipev-CO2.Rmd
more on paper figures

# monthly 
taka_month <- taka %>%
  group_by(montht) %>%
  dplyr::mutate(
    pco2_thermal =
      mean(pco2_fb_insi_temp, na.rm = TRUE) * exp(0.0423 * (
        temp_insitu_11m - mean(temp_insitu_11m, na.rm = TRUE)
      )),
    pco2_nonthermal = pco2_fb_insi_temp * exp(0.0423 * (
      mean(temp_insitu_11m, na.rm = TRUE) - temp_insitu_11m
    ))
  ) %>%
  dplyr::summarise(
    thermal = max(pco2_thermal, na.rm = TRUE) - min(pco2_thermal, na.rm = TRUE),
    nonthermal = max(pco2_nonthermal, na.rm = TRUE) - min(pco2_nonthermal, na.rm = TRUE),
    thermal_nonthermal = thermal / nonthermal
  ) %>%
  ungroup()
# monthly plot
taka_month_plt <- taka_month %>% 
  ggplot() +
  geom_point(aes(x = montht, y = thermal_nonthermal), col="blue", size = 2) +
  labs(x="Month", y= "Thermal/non-thermal control of pCO2")  +
  geom_hline(yintercept = 1, linetype = "dashed") +  
  theme_bw() + Mytheme(size_labs = 12) + theme(aspect.ratio = 1) +
  coord_fixed(ratio = 1, expand = TRUE, clip = "on")
taka_month_plt
ggsave(filename = "figures/taka_month.png", plot = taka_month_plt, width = 25, height = 25, units = "cm")

@@

```{r, fig.width=10, fig.height=10, echo=FALSE, message= FALSE}
size_point <- 0.3 # size of data points
alpha <- 1 #transparency
size_labs <- 6 # font size labels

ts_sal <- d_all %>%
  ggplot(aes(x = datetime, y = sal_fb_mix), na.rm = TRUE) +
  scale_x_datetime(breaks = date_breaks("1 year"), labels = date_format("%Y")) +
  geom_point(colour="blue", alpha = alpha, na.rm = TRUE, size = size_point) +
  labs(title="Salinity (below 8 m)",x="", y="Salinity") +
  Mytheme(size_labs = size_labs)

# temp_insitu_11m
ts_temp <- d_all %>%
  ggplot(aes(x = datetime, y = temp_insitu_11m), na.rm = TRUE) +
  scale_x_datetime(breaks = date_breaks("1 year"), labels = date_format("%Y")) +
  geom_point(colour="blue", alpha = alpha, na.rm = TRUE, size = size_point) +
  labs(title="In situ temperature (11 m)",x="", y="Temperature (°C)") +
  Mytheme(size_labs = size_labs)

# PAR
ts_par_air <- d_all %>%
  ggplot(aes(x = datetime, y = par_air), na.rm = TRUE) +
  scale_x_datetime(breaks = date_breaks("1 year"), labels = date_format("%Y")) +
  geom_point(colour="blue", alpha = alpha, na.rm = TRUE, size = size_point) +
  labs(title="Atmospheric PAR",x="", y="PAR (umol m-2 s-1)") +
  Mytheme(size_labs = size_labs)

# PCO2 
ts_pCO2 <- d_all %>%
  ggplot() +
  scale_x_datetime(breaks = date_breaks("1 year"), labels = date_format("%Y")) +
  geom_point(aes(x = datetime, y = pco2_fb_insi_temp), 
             colour="blue", alpha = alpha, na.rm = TRUE, size = size_point) +
  geom_point(aes(x = datetime, y = pco2_calc_fb_insi_temp), colour="red", alpha = alpha, na.rm = TRUE, size = size_point*3) + 
  labs(title=expression(paste(pCO[2], " at 11 m and in situ temperature")),
       subtitle = "measured in blue; calculated from CT and AT in red",
       x="", y=TeX("pCO$_2$ ($\\mu atm$)")) +
  Mytheme(size_labs = size_labs)

# SEAFET pH
ts_pH <- d_all %>% 
  dplyr::filter(depth > 8) %>% 
  ggplot() +
  scale_x_datetime(breaks = date_breaks("1 year"), labels = date_format("%Y")) +
  geom_point(aes(x = datetime, y = pHint_tot_sf), 
             colour="blue", alpha = alpha, na.rm = TRUE, size = size_point) + 
  geom_point(aes(x = datetime, y = ph_s_sf_temp_insi), colour="red", alpha = alpha, na.rm = TRUE, size = size_point*3) + 
  labs(title="In situ pH below 8 m at in situ temperature", 
       subtitle = "seaFET in blue; spectrophotometric in red",
       x = "", y=expression(paste(pH[T]))) +
  Mytheme(size_labs = size_labs) +
  theme(legend.position = c(0.1, 0.5),
        legend.key.height = unit(0.5, 'cm'))

# AT
ts_at <- d_all %>%
  ggplot(aes(x = datetime, y = at_calc), na.rm = TRUE) +
  scale_x_datetime(breaks = date_breaks("1 year"), labels = date_format("%Y")) +
  geom_point(colour="blue", alpha = alpha, na.rm = TRUE, size = size_point) +
  geom_point(aes(x = datetime, y = at), colour="red", alpha = alpha, na.rm = TRUE, size = size_point*3) + 
  labs(title="Total alkalinity", 
       subtitle = "calculated from salinity in blue; discrete measurements in red",
       x="", y=TeX("$A_T$ ($\\mu mol\\phantom{x}kg^{-1})$")) +
  Mytheme(size_labs = size_labs)

# Combine plots
g1 <- (ts_sal / ts_temp / ts_pCO2  / ts_pH / ts_at)
g1 <- g1 + plot_annotation(tag_levels = 'A')
ggsave(filename = "figures/time_series.png", plot = g1, width = 15, height = 30, units = "cm")
#rm(list = ls(pattern = "ts_")) #delete big ggplot objects
#gc() #garbage collection
```

@@

```{r, fig.width=10, fig.height=10, echo=FALSE, message= FALSE}
size_point <- 0.5 # size of data points
alpha <- 1 #transparency
d_bp <- d_all %>% # just for this chunk
  dplyr::select(datetime, sal_fb_mix, temp_insitu_11m, pco2_fb, pHint_tot_sf, at_calc)

@@

# identify extremes
d_bp <- d_bp %>%
  dplyr::mutate(
    ext_sal_fb_mix = ifelse(
      sal_fb_mix < dec_sal_fb_mix[1] |
        sal_fb_mix > dec_sal_fb_mix[2],
      sal_fb_mix,
      NA
    ),
    ext_temp_insitu_11m = ifelse(
      temp_insitu_11m < dec_temp_insitu_11m[1] |
        temp_insitu_11m > dec_temp_insitu_11m[2],
      temp_insitu_11m,
      NA
    ),
    ext_pco2_fb = ifelse(pco2_fb < dec_pco2_fb[1] |
                           pco2_fb > dec_pco2_fb[2], pco2_fb, NA),
    ext_pHint_tot_sf = ifelse(
      pHint_tot_sf < dec_pHint_tot_sf[1] |
        pHint_tot_sf > dec_pHint_tot_sf[2],
      pHint_tot_sf,
      NA
    ),
    ext_at_calc = ifelse(at_calc < dec_at_calc[1] |
                           at_calc > dec_at_calc[2], at_calc, NA)
  )
ts_sal_bp <- d_bp %>%
  ggplot(aes(x = month(x = datetime, label = TRUE, abbr = TRUE), y = sal_fb_mix)) +
  geom_boxplot(notch=FALSE, size=0.2, fill="blue", outlier.shape = NA, na.rm = TRUE) +
  geom_point(aes(x = month(x = datetime, label = TRUE, abbr = TRUE), y = ext_sal_fb_mix), col = "lightblue", size = size_point, na.rm = TRUE) +
  labs(title="Salinity (< 8 m)", y = "Salinity", x=NULL) +
  Mytheme() +
  theme(panel.grid.major.x = element_blank(),
        aspect.ratio=0.5)
ts_temp_bp <- d_bp %>%
  ggplot(aes(x = month(x = datetime, label = TRUE, abbr = TRUE), y = temp_insitu_11m)) +
  geom_boxplot(notch=FALSE, size=0.2, fill="blue", outlier.shape = NA, na.rm = TRUE) +
  geom_point(aes(x = month(x = datetime, label = TRUE, abbr = TRUE), y = ext_temp_insitu_11m), col = "lightblue", size = size_point, na.rm = TRUE) +
  labs(title="In situ temperature (11 m)", y = "Temperature (°C)", x=NULL) +
  Mytheme() +
  theme(panel.grid.major.x = element_blank(),
        aspect.ratio=0.5)
ts_pco2_bp <- d_bp %>%
  ggplot(aes(x = month(x = datetime, label = TRUE, abbr = TRUE), y = pco2_fb)) +
  geom_boxplot(notch=FALSE, size=0.2, fill="blue", outlier.shape = NA, na.rm = TRUE) +
  geom_point(aes(x = month(x = datetime, label = TRUE, abbr = TRUE), y = ext_pco2_fb), col = "lightblue", size = size_point, na.rm = TRUE) +
  labs(title=expression(paste(pCO[2], " at 11 m and in situ temperature")), y = TeX("pCO$_2$ ($\\mu atm$)"), x=NULL) +
  Mytheme() +
  theme(panel.grid.major.x = element_blank(),
        aspect.ratio=0.5)
ts_pH_bp <- d_bp %>%
  ggplot(aes(x = month(x = datetime, label = TRUE, abbr = TRUE), y = pHint_tot_sf)) +
  geom_boxplot(notch=FALSE, size=0.2, fill="blue", outlier.shape = NA, na.rm = TRUE) +
  geom_point(aes(x = month(x = datetime, label = TRUE, abbr = TRUE), y = ext_pHint_tot_sf), col = "lightblue", size = size_point, na.rm = TRUE) +
  labs(title="In situ pH below 8 m at in situ temperature", y = expression(paste(pH[T])), x=NULL) +
  Mytheme() +
  theme(panel.grid.major.x = element_blank(),
        aspect.ratio=0.5)
ts_at_bp <- d_bp %>%
  ggplot(aes(x = month(x = datetime, label = TRUE, abbr = TRUE), y = at_calc)) +
  geom_boxplot(notch=FALSE, size=0.2, fill="blue", outlier.shape = NA, na.rm = TRUE) +
  geom_point(aes(x = month(x = datetime, label = TRUE, abbr = TRUE), y = ext_at_calc), col = "lightblue", size = size_point, na.rm = TRUE) +
  labs(title="Total alkalinity calculated from salinity", y = TeX("$A_T$ ($\\mu mol\\phantom{x}kg^{-1})$"), x=NULL) +
  Mytheme() +
  theme(panel.grid.major.x = element_blank(),
        aspect.ratio=0.5)
# Combine plots
g2 <- (ts_sal_bp / ts_temp_bp / ts_pco2_bp / ts_pH_bp / ts_at_bp)
g2 <- g2 + plot_annotation(tag_levels = 'A')
ggsave(filename = "figures/ts_boxplots.png", plot = g2)
```

@@

```{r, echo=FALSE, message= FALSE, warning=FALSE}
g <- (ts_sal / ts_temp / ts_pCO2  / ts_pH / ts_at) |
  (ts_sal_bp / ts_temp_bp / ts_pco2_bp / ts_pH_bp / ts_at_bp)
g <- g + plot_annotation(tag_levels = 'A')
ggsave(filename = "figures/ts_ts_boxplots.png", plot = g)
rm(list = ls(pattern = "ts_"), d_bp) #delete big ggplot objects and temp data
#gc() #garbage collection
```

>>>>

# monthly 
taka_month <- taka %>%
  group_by(montht) %>%
  dplyr::mutate(
    pco2_thermal =
      mean(pco2_fb_insi_temp, na.rm = TRUE) * exp(0.0423 * (
        temp_insitu_11m - mean(temp_insitu_11m, na.rm = TRUE)
      )),
    pco2_nonthermal = pco2_fb_insi_temp * exp(0.0423 * (
      mean(temp_insitu_11m, na.rm = TRUE) - temp_insitu_11m
    ))
  ) %>%
  dplyr::summarise(
    thermal = max(pco2_thermal, na.rm = TRUE) - min(pco2_thermal, na.rm = TRUE),
    nonthermal = max(pco2_nonthermal, na.rm = TRUE) - min(pco2_nonthermal, na.rm = TRUE),
    thermal_nonthermal = thermal / nonthermal
  ) %>%
  ungroup()
# monthly plot
taka_month_plt <- taka_month %>% 
  ggplot() +
  geom_point(aes(x = montht, y = thermal_nonthermal), col="blue", size = 2) +
  labs(x="Month", y= latex2exp::TeX("Thermal/non-thermal control of $pCO_{2}"))  +
  geom_hline(yintercept = 1, linetype = "dashed") +  
  theme_bw() + Mytheme(size_labs = 12) + theme(aspect.ratio = 1) +
  coord_fixed(ratio = 1, expand = TRUE, clip = "on")
taka_month_plt
ggsave(filename = "figures/taka_month.png", plot = taka_month_plt, width = 25, height = 25, units = "cm")
ggsave(filename = "figures/essd/fig05.pdf", plot = taka_month_plt, width = 25, height = 25, units = "cm")

@@

```{r, fig.width=10, fig.height=10, echo=FALSE, message= FALSE}
# legend:
# - Salinity below 8m
# - In situ temperature (11m)
# - pCO2 at 11 m and insitu Temperature, measued in blue, calculated from CT an AT in red
# - In situ pH below 8 m at in situ Temperature, seaFET in blue, spectrophotometric in red
# - Total alkalinity calculated from salinity in blue, discrete measurement in red

size_point <- 0.3 # size of data points
alpha <- 1 #transparency
size_labs <- 12 # font size labels

ts_sal <- d_all %>%
  ggplot(aes(x = datetime, y = sal_fb_mix), na.rm = TRUE) +
  scale_x_datetime(breaks = date_breaks("1 year"), labels = date_format("%Y")) +
  geom_point(colour="blue", alpha = alpha, na.rm = TRUE, size = size_point) +
  labs(title=NULL,x="", y="Salinity") +
  Mytheme(size_labs = size_labs)

# temp_insitu_11m
ts_temp <- d_all %>%
  ggplot(aes(x = datetime, y = temp_insitu_11m), na.rm = TRUE) +
  scale_x_datetime(breaks = date_breaks("1 year"), labels = date_format("%Y")) +
  geom_point(colour="blue", alpha = alpha, na.rm = TRUE, size = size_point) +
  labs(title=NULL,x="", y="Temperature (°C)") +
  Mytheme(size_labs = size_labs)

# PAR
ts_par_air <- d_all %>%
  ggplot(aes(x = datetime, y = par_air), na.rm = TRUE) +
  scale_x_datetime(breaks = date_breaks("1 year"), labels = date_format("%Y")) +
  geom_point(colour="blue", alpha = alpha, na.rm = TRUE, size = size_point) +
  labs(title="Atmospheric PAR",x="", 
       y=latex2exp::TeX("PAR ($\\mu mol\\phantom{x}m^{-2}\\phantom{x}s^{-1})$")) +
  Mytheme(size_labs = size_labs)

# PCO2 
ts_pCO2 <- d_all %>%
  ggplot() +
  scale_x_datetime(breaks = date_breaks("1 year"), labels = date_format("%Y")) +
  geom_point(aes(x = datetime, y = pco2_fb_insi_temp), 
             colour="blue", alpha = alpha, na.rm = TRUE, size = size_point) +
  geom_point(aes(x = datetime, y = pco2_calc_fb_insi_temp), colour="red", alpha = alpha, na.rm = TRUE, size = size_point*3) + 
  labs(title=NULL,
       subtitle = NULL,
       x="", y=TeX("pCO$_2$ ($\\mu atm$)")) +
  Mytheme(size_labs = size_labs)

# SEAFET pH
ts_pH <- d_all %>% 
  dplyr::filter(depth > 8) %>% 
  ggplot() +
  scale_x_datetime(breaks = date_breaks("1 year"), labels = date_format("%Y")) +
  geom_point(aes(x = datetime, y = pHint_tot_sf), 
             colour="blue", alpha = alpha, na.rm = TRUE, size = size_point) + 
  geom_point(aes(x = datetime, y = ph_s_sf_temp_insi), colour="red", alpha = alpha, na.rm = TRUE, size = size_point*3) + 
  labs(title=NULL, 
       subtitle =NULL,
       x = "", y=expression(paste(pH[T]))) +
  Mytheme(size_labs = size_labs) +
  theme(legend.position = c(0.1, 0.5),
        legend.key.height = unit(0.5, 'cm'))

# AT
ts_at <- d_all %>%
  ggplot(aes(x = datetime, y = at_calc), na.rm = TRUE) +
  scale_x_datetime(breaks = date_breaks("1 year"), labels = date_format("%Y")) +
  geom_point(colour="blue", alpha = alpha, na.rm = TRUE, size = size_point) +
  geom_point(aes(x = datetime, y = at), colour="red", alpha = alpha, na.rm = TRUE, size = size_point*3) + 
  labs(title=NULL, 
       subtitle = NULL,
       x="", y=TeX("$A_T$ ($\\mu mol\\phantom{x}kg^{-1})$")) +
  Mytheme(size_labs = size_labs)

# Combine plots
g1 <- (ts_sal / ts_temp / ts_pCO2  / ts_pH / ts_at)
g1 <- g1 + plot_annotation(tag_levels = 'A' & theme(plot.tag = element_text(face = "bold")))
ggsave(filename = "figures/time_series.png", plot = g1, width = 15, height = 30, units = "cm")

#rm(list = ls(pattern = "ts_")) #delete big ggplot objects
#gc() #garbage collection
```

@@

```{r, fig.width=10, fig.height=10, echo=FALSE, message= FALSE}
size_point <- 0.5 # size of data points
alpha <- 1 #transparency
size_labs <- 12 # font size labels
d_bp <- d_all %>% # just for this chunk
  dplyr::select(datetime, sal_fb_mix, temp_insitu_11m, pco2_fb, pHint_tot_sf, at_calc)

@@

# identify extremes
d_bp <- d_bp %>%
  dplyr::mutate(
    ext_sal_fb_mix = ifelse(
      sal_fb_mix < dec_sal_fb_mix[1] |
        sal_fb_mix > dec_sal_fb_mix[2],
      sal_fb_mix,
      NA
    ),
    ext_temp_insitu_11m = ifelse(
      temp_insitu_11m < dec_temp_insitu_11m[1] |
        temp_insitu_11m > dec_temp_insitu_11m[2],
      temp_insitu_11m,
      NA
    ),
    ext_pco2_fb = ifelse(pco2_fb < dec_pco2_fb[1] |
                           pco2_fb > dec_pco2_fb[2], pco2_fb, NA),
    ext_pHint_tot_sf = ifelse(
      pHint_tot_sf < dec_pHint_tot_sf[1] |
        pHint_tot_sf > dec_pHint_tot_sf[2],
      pHint_tot_sf,
      NA
    ),
    ext_at_calc = ifelse(at_calc < dec_at_calc[1] |
                           at_calc > dec_at_calc[2], at_calc, NA)
  )
ts_sal_bp <- d_bp %>%
  ggplot(aes(x = month(x = datetime, label = TRUE, abbr = TRUE), y = sal_fb_mix)) +
  geom_boxplot(notch=FALSE, size=0.2, fill="blue", outlier.shape = NA, na.rm = TRUE) +
  geom_point(aes(x = month(x = datetime, label = TRUE, abbr = TRUE), y = ext_sal_fb_mix), col = "lightblue", size = size_point, na.rm = TRUE) +
  labs(title="", y = "Salinity", x=NULL) +
  Mytheme(size_labs = size_labs) +
  theme(panel.grid.major.x = element_blank(),
        aspect.ratio=0.5)
ts_temp_bp <- d_bp %>%
  ggplot(aes(x = month(x = datetime, label = TRUE, abbr = TRUE), y = temp_insitu_11m)) +
  geom_boxplot(notch=FALSE, size=0.2, fill="blue", outlier.shape = NA, na.rm = TRUE) +
  geom_point(aes(x = month(x = datetime, label = TRUE, abbr = TRUE), y = ext_temp_insitu_11m), col = "lightblue", size = size_point, na.rm = TRUE) +
  labs(title=NULL, y = "Temperature (°C)", x=NULL) +
  Mytheme(size_labs = size_labs) +
  theme(panel.grid.major.x = element_blank(),
        aspect.ratio=0.5)
ts_pco2_bp <- d_bp %>%
  ggplot(aes(x = month(x = datetime, label = TRUE, abbr = TRUE), y = pco2_fb)) +
  geom_boxplot(notch=FALSE, size=0.2, fill="blue", outlier.shape = NA, na.rm = TRUE) +
  geom_point(aes(x = month(x = datetime, label = TRUE, abbr = TRUE), y = ext_pco2_fb), col = "lightblue", size = size_point, na.rm = TRUE) +
  labs(title=NULL, y = TeX("pCO$_2$ ($\\mu atm$)"), x=NULL) +
  Mytheme(size_labs = size_labs)+
  theme(panel.grid.major.x = element_blank(),
        aspect.ratio=0.5)
ts_pH_bp <- d_bp %>%
  ggplot(aes(x = month(x = datetime, label = TRUE, abbr = TRUE), y = pHint_tot_sf)) +
  geom_boxplot(notch=FALSE, size=0.2, fill="blue", outlier.shape = NA, na.rm = TRUE) +
  geom_point(aes(x = month(x = datetime, label = TRUE, abbr = TRUE), y = ext_pHint_tot_sf), col = "lightblue", size = size_point, na.rm = TRUE) +
  labs(title=NULL, y = TeX("$pH_T$"), x=NULL) +
  Mytheme(size_labs = size_labs) +
  theme(panel.grid.major.x = element_blank(),
        aspect.ratio=0.5)
ts_at_bp <- d_bp %>%
  ggplot(aes(x = month(x = datetime, label = TRUE, abbr = TRUE), y = at_calc)) +
  geom_boxplot(notch=FALSE, size=0.2, fill="blue", outlier.shape = NA, na.rm = TRUE) +
  geom_point(aes(x = month(x = datetime, label = TRUE, abbr = TRUE), y = ext_at_calc), col = "lightblue", size = size_point, na.rm = TRUE) +
  labs(title=NULL, y = TeX("$A_T$ ($\\mu mol\\phantom{x}kg^{-1})$"), x=NULL) +
  Mytheme(size_labs = size_labs) +
  theme(panel.grid.major.x = element_blank(),
        aspect.ratio=0.5)
# Combine plots
g2 <- (ts_sal_bp / ts_temp_bp / ts_pco2_bp / ts_pH_bp / ts_at_bp)
g2 <- g2 + plot_annotation(tag_levels = 'A' & theme(plot.tag = element_text(face = "bold")))
ggsave(filename = "figures/ts_boxplots.png", plot = g2)
```

@@

```{r, echo=FALSE, message= FALSE, warning=FALSE}
g <- (ts_sal / ts_temp / ts_pCO2  / ts_pH / ts_at) |
  (ts_sal_bp / ts_temp_bp / ts_pco2_bp / ts_pH_bp / ts_at_bp)
g <- g + plot_annotation(tag_levels = 'A' & theme(plot.tag = element_text(face = "bold")))
ggsave(filename = "figures/essd/fig06.pdf", plot = g)
# legend:
# - Salinity below 8m
# - In situ temperature (11m)
# - pCO2 at 11 m and insitu Temperature, measued in blue, calculated from CT an AT in red
# - In situ pH below 8 m at in situ Temperature, seaFET in blue, spectrophotometric in red
# - Total alkalinity calculated from salinity in blue, discrete measurement in red

#rm(list = ls(pattern = "ts_"), d_bp) #delete big ggplot objects and temp data
#gc() #garbage collection
```
